---
title: "Death Caused By Drug Use"
author: "Bryan Lin, Alex Chau, Winnie Hou, Kathy Lam"
date: "5/24/2019"
output: html_document
---

```{r setup, include=FALSE}
accidental_drug_deaths <- read.csv("data/Accidental_Drug_Related_Deaths_2012-2018.csv", stringsAsFactors = FALSE)
source("scripts/summary-info.R")
source("scripts/aggregate-table.R")
source("scripts/accidental_drug_related_deaths_plot.R")
source("scripts/drug_use_by_age_data_plot.R")
source("scripts/Drug_Overdose_Death.R")
library("knitr")
```

## Introduction
This project's goal is to see if there is a correlation between drug use and deaths. The data that we are using were obtained from [Data World](https://data.world/fivethirtyeight/drug-use-by-age), [Data.gov](https://catalog.data.gov/dataset/accidental-drug-related-deaths-january-2012-sept-2015), and the [CDC](https://www.cdc.gov/nchs/nvss/vsrr/drug-overdose-data.htm). Some questions that we are trying to answer is *if drug use increases as age increases* or *if the number of deaths have increased over the years for the younger generation*. Drug use has increased over the years, so by answering some of these questions, we believe we can adjust the healthcare approach of **addressing** and **preventing** further deaths by drug use. 

## Summary Information
One of the datasets that we are using comes from [Data.gov](https://catalog.data.gov/dataset/accidental-drug-related-deaths-january-2012-sept-2015) which gives data about accidental drug related deaths from 2012 to 2018. The dataset has a length of `r get_summary_info(accidental_drug_deaths)$length`. From this dataset, it shows that the average age of those who died of drug use is `r get_summary_info(accidental_drug_deaths)$average_age_of_drug_used_death` years old. If we look at the percent of males who died of drug use in comparison to the percent of females, it shows that the percent of males was `r get_summary_info(accidental_drug_deaths)$male_percentage`% versus the percent of females was `r get_summary_info(accidental_drug_deaths)$female_percentage`%. Interestingly, the race that has the highest amounts of deaths due to drug use is `r get_summary_info(accidental_drug_deaths)$most_race`. Lastly, based from the percentages and the dataset, the sex that has the most deaths due to drug use is `r get_summary_info(accidental_drug_deaths)$most_sex`.

## Monthly Breakdown of Deaths Due To Opioids
Here are the number of deaths that have occurred just in 2018 from Janurary to September due to the use of opiods:

```{r drug_od_table, echo = FALSE}
drug_od_table(drug_od_df)
```

The reason we included the table was to answer the question of *what is the monthly breakdown of the type of drug use and the number of deaths?*. We wanted to see just in 2018 alone, based on each month, how many people have died from drug overdose. By having this table, it gives a clear layout where we are able to analyze if there is a correlation between the number of deaths and the month due to the use of opioids. Based on the table, surprisingly, from January to September, the number of deaths are quite close to each other. Yet, there is a small miniscule difference where January has the highest amount of deaths while September has the lowest. This shows that a specific month could play a factor in the amount of people who tend to use opioiods. If we think about this in terms of seasons, it kind of shows as the season gets better, the number of deaths decrease. January-March is winter and the most gloomy time of the year which could result in higher number of deaths, while June-September is summer which could result in slightly lower deaths.

## Drug Use By Age
Here is a histogram that shows the use of drugs by age:

```{r drug_age_plot, echo = FALSE}
drug_age_plot(drug_age_data)
```

The reason why we included this chart was to see the relationship between age and drugs. We wanted to know if there is a correlation between the two and how this could be linked to deaths at a certain age due to drug overdose. By including a histogram, it allows us to easily analyze the different ages and the number of people of that age group uses drugs. From this table, it shows that the highest age group that uses the most drugs are from ages 35-49. This is surprising as we would assume teenagers would be the highest, but from the table, it shows that ages 22-25 are the second highest when it comes to drug use. Another information that was revealed that was surprising was that from ages 12-17, there are more number of users of drugs than from ages 18-21.

## Deaths By Drug Overdose From 2015-2018
Here is a pie chart that shows the number of drug overdose deaths by state:

```{r plot_drug_overdose, echo = FALSE}
plot_drug_overdose(drug_overdose)
```

The reason why we included this map was to see the relationship between a specific area and drug overdose deaths. We wanted to know if certain states had higher drug overdose deaths and analyze to see if there could be a factor in play. By using a map that is color-coated and can be hovered over for more information, it allows us to visually see which states have higher death counts just based on the darkness of the color. From this data, it shows that the state with the highest number of deaths due to drug overdose is Oklahoma. The state with the lowest number of deaths due to drug overdose is Missouri with only 9155 deaths. By looking at these states, we can think about whether different factors could be at play in the number of deaths by drug overdose. For example, political laws could be a factor that could affect the use of the drugs in that state.

## Place of Death from Drug Overdose in Connecticut
Here is a map that shows the different places where people died from drug overdose in Connecticut:

```{r plot_accident_drug, echo = FALSE}
plot_accident_drug(accident_drug)
```

The reason we included this map was to show a specific state, specifically Connecticut and how race, sex and the cause of death from drug use are related. By having the map, we are able to see different points of which counties have more concentrated points, which means more deaths and by clicking on the points, we are able to get information about the specific person. But, also by including the color of the different races, we are able to visiualize which race had the most deaths just in the state of Connecticut. From this data, it reveals that the majority of the deaths in Connecticut are white people.









